Density dependence

### 2020-10-28 R/competition wrangling.Rmd
revised row and column locations and re-ran density dependence analysis.

```{r}
ha_density <-
  ha_xy %>% 
  group_by(plot, year) %>% 
  group_split() %>% 
  # head() %>% #test with 6 df to see how long it will take.
  future_map(~{
    add_density_cols(.x, L = 0, U = 5, by = 0.5)
  }) %>% 
  future_map(~{
    add_density_cols(.x, L = 0, U = 3, by = 3)
  }) %>% 
  bind_rows()
beepr::beep(4)
```

>>>>

```{r}
ha_density <-
  ha_xy %>% 
  group_by(plot, year) %>% 
  group_split() %>% 
  # head() %>% #test with 6 df to see how long it will take.
  future_map(~{
    add_density_cols(.x, L = 0, U = 5, by = 0.5)
  }) %>% 
  future_map(~{
    add_density_cols(.x, L = 0, U = 1, by = 1) #first meter seems most important
  }) %>% 
  bind_rows()
beepr::beep(4)
```

### 2020-10-28 R/vital rate models.Rmd
revised row and column locations and re-ran density dependence analysis.

```{r}
s_df %>% 
  ggplot(aes(x = exp(x), y = plogis(fit + coef(s_post_cf)[1]))) +
  geom_line() +
  geom_ribbon(aes(ymin = plogis(fit + coef(s_post_cf)[1] - se),
                  ymax = plogis(fit + coef(s_post_cf)[1] + se)),
              alpha = 0.2) +
  labs(x = "size (#shoots)", y = "P(survival)") +
  facet_wrap(~habitat) +
  stat_summary_bin(data = demog_post %>% filter(habitat %in% c("1-ha", "CF")),
                   geom = "point",
                   fun = mean,
                   bins = 7,
                   aes(y = surv, x = shts)) +
  theme_bw() +
  plot_annotation(caption = "Note: these are from two separate models, not a size–habitat interaction!")
```

>>>>

```{r}
s_df %>% 
  ggplot(aes(x = exp(x), y = plogis(fit + coef(s_post_cf)[1]))) +
  geom_line() +
  geom_ribbon(aes(ymin = plogis(fit + coef(s_post_cf)[1] - se),
                  ymax = plogis(fit + coef(s_post_cf)[1] + se)),
              alpha = 0.2) +
  labs(x = "size (#shoots)", y = "P(survival)") +
  facet_wrap(~habitat) +
  stat_summary_bin(data = demog_post %>% filter(habitat %in% c("1-ha", "CF")),
                   geom = "point",
                   fun = mean,
                   bins = 20,
                   aes(y = surv, x = shts)) +
  theme_bw() +
  plot_annotation(caption = "Note: these are from two separate models, not a size–habitat interaction!")

demog_full %>% ggplot(aes(shts)) + geom_histogram(bins = 20)
hist(demog_full$shts)
```
```{r}
demog_full %>% 
  filter(shts >= 15)
demog_full %>% 
  filter(ha_id_number == 5310) %>% View()
```

### 2020-10-28 analysis/paper/appendix_density_dependence.Rmd
revised row and column locations and re-ran density dependence analysis.

-   Remove saplings and update stats

@@

```{r eval=FALSE, include=FALSE}
g_growth_pois <- update(g_growth, family = poisson(link = "log"))
g_growth_norm <- update(g_growth, family = gaussian(link = "log")) #residuals are leptokurtic with gaussian
AICtab(g_growth, g_growth_pois, g_growth_scat)
```

@@

Plants had a greater number of shoots in year t+1 as conspecific density increased, all else being average. The effects of conspecific density were strongest around 1.5 -- 2 meters from the focal plant (Figure \@ref(fig:growth-cb)).

```{r growth-slice, include=TRUE, fig.cap="this is a figure"}
pred_growth %>%
  filter(min_dist < 0.1) %>%
  filter(nearest(lag, 2)) %>% 
  ggplot(aes(x = x, y = exp(fitted), ymin = exp(fitted - 1.96*se.fit), ymax = exp(fitted + 1.96*se.fit))) +
  geom_jitter(data = ha %>%
                filter(across(shts_0.5_1:shts_4.5_5, ~!is.na(.))),
              aes(x = shts_1.5_2, y = shts_next),
              alpha = 0.2,
              inherit.aes = FALSE) +
  geom_line(color = "blue") + 
  geom_ribbon(alpha = 0.4) +

  labs(title = "Near 2m from focal plant", 
       x = "conspecific density (shoots/m^2)",
       y = "#shoots (t+1)") +
  theme_bw()
```

@@

Plant size had a significant, nearly quadratic (on a logit scale), effect on survival in continuous forest ($\chi^2$ = 112.2, edf = 2.1, P \< 0.001), but a linear (on a logit scale) effect in 1 ha fragments ($\chi^2$ = 15.63, edf = 1, P \< 0.001), but their 95% confidence intervals overlap (Figure \@ref(fig:surv-plot)). However, there were no significant effects of habitat ($\chi^2$ = 0.002, df = 1, P = 0.964) or conspecific density ($\chi^2$ = 5.326, edf = 2, P = 0.0699) on survival.

@@

```{r eval=FALSE, include=FALSE}
pred_growth %>%
  filter(min_dist < 0.1) %>% 
  filter(nearest(x, 5)) %>% 
  ggplot(aes(x = lag, y = exp(fitted), ymin = exp(fitted - 1.96*se.fit), ymax = exp(fitted + 1.96*se.fit))) +
  # geom_jitter(data = ha %>%
  #               filter(across(shts_0.5_1:shts_4.5_5, ~!is.na(.))),
  #             aes(x = shts_1.5_2, y = shts_next),
  #             alpha = 0.2,
  #             inherit.aes = FALSE) +
  # geom_smooth(color = "blue", se = FALSE) + 
  geom_line(color = "blue") +
  geom_ribbon(alpha = 0.4) +

  labs(title = "Conspecific density of 5 shoots/m^2", 
       x = "Midpoint distance from focal plant (m)",
       y = "#shoots (t+1)") +
  theme_bw()
```

```{r}
mnorm <-
  glmer(
    shts_next ~ 
      log_shts *
      habitat *
      shts_0_3 +
      (1|ha_id_number) +
      (1|plot) +
      (1|year_post),
    family = gaussian(link = "log"), #can't use scat() in glmer()
    data = ha
  )
# Random effects
mnorm2 <- update(mnorm, . ~ . - (1|ha_id_number))
AICtab(mnorm, mnorm2) #keep ID number
mnorm3 <- update(mnorm, .~. - (1|plot)) 
AICtab(mnorm, mnorm2, mnorm3) #remove plot
mnorm4 <- update(mnorm3, .~.-(1|year_post)) #remove year
AICtab(mnorm, mnorm2, mnorm3, mnorm4)

# Interactions
mnorm5 <- update(mnorm4, .~. - log_shts:habitat:shts_0_3)
AICtab(mnorm4, mnorm5) #good to remove 3-way interaction
mnorm6 <- update(mnorm5, .~. -log_shts:habitat)
mnorm7 <- update(mnorm5, .~. -log_shts:shts_0_3)
mnorm8 <- update(mnorm5, .~. -habitat:shts_0_3)
mnorm9 <- update(mnorm5, .~. -habitat:shts_0_3 -log_shts:shts_0_3)
mnorm10 <- update(mnorm5, .~. -habitat:shts_0_3 -log_shts:habitat)
mnorm11 <- update(mnorm5, .~. -habitat:shts_0_3 -log_shts:habitat -log_shts:shts_0_3)
AICtab(mnorm5, mnorm6, mnorm7, mnorm8, mnorm9, mnorm10, mnorm11) #good to remove all interactions

Anova(mnorm11)
fixef(mnorm11)
```

@@

@@

ha_sub <- ha %>% filter(!is.na(shts_0_3), !is.na(log_shts))
newdata <-
  tibble(log_shts = mean(ha_sub$log_shts),
         habitat = "1-ha", 
         ha_id_number = "new",
         shts_0_3 = seq(min(ha_sub$shts_0_3, na.rm = TRUE),
                          max(ha_sub$shts_0_3, na.rm = TRUE),
                          length.out = 50))

augment(mnorm11, newdata = newdata, re.form = NA) %>% 
  ggplot(aes(x = shts_0_3, y = exp(.fitted))) +
  geom_jitter(data = ha_sub, aes(x = shts_0_3, y = shts_next), alpha = 0.2) +
  geom_line(color = "blue") +
  # geom_ribbon() + #don't have CI yet.  Need to do something like bootMer to get them with bootstrapping.
  labs(x = "shoot density within 3m (shoots/m^2)", y = "shoots (t+1)") +
  theme_bw()
```

>>>>

-   Update stats in writing

@@

```{r eval=FALSE, include=FALSE}
g_growth_pois <- update(g_growth_scat, family = poisson(link = "log"))
g_growth_norm <- update(g_growth_scat, family = gaussian(link = "log")) #residuals are leptokurtic with gaussian
AICtab(g_growth_norm, g_growth_pois, g_growth_scat)
```

@@

Plants had a greater number of shoots in year t+1 as conspecific density increased, all else being average. The effects of conspecific density were strongest around 0--1 meters from the focal plant (Figure \@ref(fig:growth-cb)).

```{r growth-slice, include=TRUE, fig.cap="this is a figure"}
pred_growth %>%
  filter(min_dist < 0.1) %>%
  filter(nearest(lag, 0.5)) %>% 
  ggplot(aes(x = x, y = exp(fitted), ymin = exp(fitted - 1.96*se.fit), ymax = exp(fitted + 1.96*se.fit))) +
  geom_jitter(data = ha %>%
                filter(across(shts_0.5_1:shts_4.5_5, ~!is.na(.))),
              aes(x = shts_0_0.5, y = shts_next),
              alpha = 0.2,
              inherit.aes = FALSE) +
  geom_line(color = "blue") + 
  geom_ribbon(alpha = 0.4) +

  labs(title = "Near 0.5m from focal plant", 
       x = "conspecific density (shoots/m^2)",
       y = "#shoots (t+1)") +
  theme_bw()
```

@@

Plant size had a significant, nearly quadratic (on a logit scale), effect on survival in continuous forest ($\chi^2$ = 112.2, edf = 2.1, P \< 0.001), but a linear (on a logit scale) effect in 1 ha fragments ($\chi^2$ = 15.63, edf = 1, P \< 0.001), but their 95% confidence intervals overlap (Figure \@ref(fig:surv-plot)). However, there were no significant effects of habitat ($\chi^2$ = 0.002, df = 1, P = 0.964) or conspecific density ($\chi^2$ = 5.36, edf = 2, P = 0.0688) on survival.

@@

```{r}
mnorm <-
  glmer(
    shts_next ~ 
      log_shts *
      habitat *
      shts_0_1 +
      (1|ha_id_number) +
      (1|plot) +
      (1|year_post),
    family = gaussian(link = "log"), #can't use scat() in glmer()
    data = ha
  )
# Random effects
mnorm2 <- update(mnorm, . ~ . - (1|ha_id_number))
AICtab(mnorm, mnorm2) #keep ID number
mnorm3 <- update(mnorm, .~. - (1|plot)) 
AICtab(mnorm, mnorm2, mnorm3) #remove plot
mnorm4 <- update(mnorm3, .~.-(1|year_post)) #remove year
AICtab(mnorm, mnorm2, mnorm3, mnorm4)

# Interactions
mnorm5 <- update(mnorm4, .~. - log_shts:habitat:shts_0_1)
AICtab(mnorm4, mnorm5) #good to remove 3-way interaction
mnorm6 <- update(mnorm5, .~. -log_shts:habitat)
mnorm7 <- update(mnorm5, .~. -log_shts:shts_0_1)
mnorm8 <- update(mnorm5, .~. -habitat:shts_0_1)
mnorm9 <- update(mnorm5, .~. -habitat:shts_0_1 -log_shts:shts_0_1)
mnorm10 <- update(mnorm5, .~. -habitat:shts_0_1 -log_shts:habitat)
mnorm11 <- update(mnorm5, .~. -habitat:shts_0_1 -log_shts:habitat -log_shts:shts_0_1)
AICtab(mnorm5, mnorm6, mnorm7, mnorm8, mnorm9, mnorm10, mnorm11) #good to remove all interactions except size by density

Anova(mnorm10)
fixef(mnorm10)
```

@@

marginally significant size x conspecific density interaction (p = 0.046). No significant main effect of density.

@@

ha_sub <- ha %>% filter(!is.na(shts_0_1), !is.na(log_shts))
newdata <-
  tibble(log_shts = mean(ha_sub$log_shts),
         habitat = "1-ha", 
         ha_id_number = "new",
         shts_0_1 = seq(min(ha_sub$shts_0_1, na.rm = TRUE),
                          max(ha_sub$shts_0_1, na.rm = TRUE),
                          length.out = 50))

augment(mnorm11, newdata = newdata, re.form = NA) %>% 
  ggplot(aes(x = shts_0_1, y = exp(.fitted))) +
  geom_jitter(data = ha_sub, aes(x = shts_0_1, y = shts_next), alpha = 0.2) +
  geom_line(color = "blue") +
  # geom_ribbon() + #don't have CI yet.  Need to do something like bootMer to get them with bootstrapping.
  labs(x = "shoot density within 3m (shoots/m^2)", y = "shoots (t+1)") +
  theme_bw()
```
